Home Nighttime Cough Monitoring

### 2024-11-01 Lilah Melzer/cough_monitor_analysis.qmd
Update cough_monitor_analysis.qmd

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(lubridate)
library(arsenal)
library(rspiro)
library(childsds)
library(lmerTest)
library(performance)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Import REDCap data
source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))

@@

start_end <- read.csv("./Data_Clean/Start and end dates.csv", na.strings = "")

@@

vacations <- read.csv("./Data_Clean/Vacation days.csv", na.strings = "")

@@

@@

  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",

@@

# Add start and end dates, filter

@@

agg_counts <- left_join(agg_counts, start_end)
  filter(Date >= Start & Date <= End)

@@

  filter(!Vacation) %>%

@@

@@

sick <- data %>%
  select(sid, pulm_ex_v2.factor) %>%
    antib_dt = mdy(antib_dt), 
    px_start = antib_dt - 7,
    px_end = antib_dt + 28

@@

data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
  levels = 1:2,
  labels = c("CF", "Non-CF")

@@

demo <- data %>%

@@

    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor

@@

  select(age_at_consent, height, weight, nih_sex.factor, pre_fev) %>%
  filter(nih_sex.factor %in% c("Male", "Female"))

@@

  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),

@@

  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,

@@

  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),

@@

@@

# Exclude vacation days

# Set table options for the document
mycontrols <-
  tableby.control(
    numeric.stats = c("Nmiss", "medianq1q3", "range"),
    numeric.test = "kwt"
  )
```

@@

```{r results='asis'}
#| label: tbl-mean-demo
#| tbl-cap: Participant characteristics
f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
f <- update(f, . ~ . - Diagnosis - sid - median_coughs + kwt(median_coughs))
t1 <- tableby(f, data = demo, control = mycontrols)
summary(t1, pfootnote = T)
```

@@

```{r}
mod <- glmer(Coughs ~ Night + Diagnosis + (Night | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
```

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(tidyverse)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
```

```{r data import}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "Cough Monitor 23-0382"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
```

@@

@@

start_end <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Start and end dates.csv",
  na.strings = ""

@@

vacations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Vacation days.csv",
  na.strings = ""

@@

# Make a grid of everyone's possible dates
dates <- start_end %>%
  rowwise() %>%
  transmute(sid,
    Date = list(seq(Start, End - 1, by = "day"))
  unnest(Date)

@@

  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",

@@

# Merge with all possible dates

@@

agg_counts <- left_join(dates, agg_counts)

@@

  filter(!Vacation | is.na(Vacation)) %>%

@@

agg_counts$Vacation <- NULL

@@

sick <- df %>%
  select(sid, pulm_ex_v2, pulm_ex_spec_v2) %>%
  filter(pulm_ex_v2 == "Yes")
sick$pulm_ex_abx_start <- mdy(c("4/17/24", "5/8/24"))
sick$pulm_ex_start <- sick$pulm_ex_abx_start - 7
sick$pulm_ex_end <- sick$pulm_ex_abx_start + 28
# Make a grid of everyone's sick days and merge
sick <- sick %>%
  select(sid, pulm_ex_start, pulm_ex_end) %>%
  rowwise() %>%
  transmute(sid,
    Date = list(seq(pulm_ex_start, pulm_ex_end, by = "day"))
  unnest(Date) %>%
  mutate(Sick = T)
sick$sid <- as.numeric(sick$sid)
agg_counts <- left_join(agg_counts, sick)
# Format sick column
agg_counts$Sick[is.na(agg_counts$Sick)] <- F
agg_counts$Sick <- factor(agg_counts$Sick,
  levels = c(F, T),
  labels = c("No", "Yes")

@@

df$Diagnosis <- factor(df$redcap_event_name,
  levels = c(
    "Visits (Arm 1: CF)", "Interim (Arm 1: CF)", "Visits (Arm 2: Control)",
    "Interim (Arm 2: Control)"
  ),
  labels = c("CF", "CF", "Non-CF", "Non-CF")

@@

df$Diagnosis <- relevel(df$Diagnosis, ref = "Non-CF")
demo <- df %>%

@@

    nih_sex:nih_ethnicity, mutation1, mutation2,
    cftr_mod, cftr_mod_type, mucolytic:antih

@@

  select(sid, age_at_consent, height, weight, nih_sex, pre_fev) %>%
  filter(nih_sex %in% c("Male", "Female"))

@@

  factor(fev1$nih_sex, levels = c("Male", "Female")),

@@

  age = fev1$age_at_consent, sex = fev1$nih_sex,

@@

  factor(rowSums(demo[, c("mutation1", "mutation2")] == "F508del"),

@@

# Save
write.csv(agg_counts, file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)

@@

# Set table options for the document
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct"),
    numeric.test = "kwt"
```

@@

```{r results='asis'}
#| label: tbl-mean-demo
#| tbl-cap: Participant characteristics
f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
f <- update(f, . ~ . - Diagnosis - sid)
t1 <- tableby(f, data = demo, control = mycontrols)
summary(t1, pfootnote = T)
```

# Data for ATS abstract

```{r results='asis'}
per_person <- agg_counts %>%
  group_by(sid, Diagnosis) %>%
  summarise(
    `Nights` = n(),
    `Nights With Data` = sum(!is.na(Coughs)),
    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
    `Median Coughs per Night` = median(Coughs, na.rm = T),
    .groups = "drop"
  )
t2 <- tableby(
  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
    `Median Coughs per Night`,
  data = per_person, control = mycontrols
)
summary(t2, pfootnote = T)
```

@@

```{r}
mod <- glmer(Coughs ~ Night + Diagnosis + (Night | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
# check_model(mod, re_formula = NA, residual_type = "normal")
```

### By sick status

```{r}
mod <- glmer(Coughs ~ Night + Diagnosis + Sick + (Night | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
# check_model(mod, re_formula = NA, residual_type = "normal")
```
